Cell Decompositions for Rank Two Quiver Grassmannians

### 2018-03-11 Grassmannian_cells.tex
Committing to pull

\begin{document}
\begin{abstract}
  We prove that all quiver Grassmannians for preprojective representations of a generalized Kronecker quiver admit a cell decomposition into affine spaces.  
  In the process, we introduce a class of regular representations which arise as quotients of consecutive preprojective representations.
  Cell decompositions for quiver Grassmannians of these ``truncated preprojectives'' are also established. 
  We also provide two natural combinatorial labelings for these cells.
  On the one hand, the cells are in bijection with compatible pairs in a maximal Dyck path as predicted by the theory of cluster algebras.
  On the other hand, they are labeled by certain subsets of a so-called $2$-quiver attached to a preprojective representation.
  As there is a natural bijection between these two labelings this gives also a geometric explanation for the appearance of Dyck path combinatorics in the theory of quiver Grassmannians. 
\end{abstract}

@@

The next step is to investigate conditions under which the fixed point components admit a cell decomposition.
Here the Caldero-Chapoton map comes into play.
In the case of the generalized Kronecker quiver, it turns out that a natural filtration of a fixed preprojective representation by preprojectives of smaller dimension transfers to the universal covering quiver.
These filtration can be successively described by short exact sequences.
The main advantage when passing to the universal covering is that the preprojective subrepresentations of the same dimension become orthogonal, a property which rigidifies the situation in a sense.
In the end, this machinery can be used to recursively build cell decompositions of all quiver Grassmannians of lifted (truncated) preprojective representations.
As all the quiver Grassmannians of the (non-lifted) representation are smooth, this combines with the torus action to give a cell decomposition of these quiver Grassmannians.

@@

As mentioned above, the Laurent polynomial expressions for cluster variables have been described using the representation theory of quivers \cite{cc,ck}: the cluster variables are generating functions for Euler characteristics of quiver Grassmannians.
For rank two cluster algebras, the Laurent expressions of cluster variables can also be computed using a certain Dyck path combinatorics \cite{llz}.
The confluence of these results gives rise to a combinatorial construction for the Euler characteristics and counting polynomials of certain quiver Grassmannians.
A consequence of our main result is a geometric explanation for these computations: we provide a one-to-one correspondence between the strong successor closed subsets and compatible pairs for an appropriate Dyck path which leads to a geometric explanation for the appearance of Dyck path combinatorics in the theory of quiver Grassmannians. 
\sayD{I need to include a conjecture about the dimensions of cells coming from the combinatorial construction of counting/Poincare polynomials}

>>>>

\begin{document}
\begin{abstract}
  We prove that all quiver Grassmannians for preprojective representations of a generalized Kronecker quiver admit a cell decomposition into affine spaces.  
  In the process, we introduce a class of regular representations which arise as quotients of consecutive preprojective representations.
  Cell decompositions for quiver Grassmannians of these ``truncated preprojectives'' are also established. 
  We also provide two natural combinatorial labelings for these cells.
  On the one hand, the cells are in bijection with compatible pairs in a maximal Dyck path as predicted by the theory of cluster algebras.
  On the other hand, they are labeled by certain subsets of a so-called $2$-quiver attached to a (truncated) preprojective representation.
  As there is a natural bijection between these two labelings this gives also a geometric explanation for the appearance of Dyck path combinatorics in the theory of quiver Grassmannians. 
\end{abstract}

@@

The next step is to investigate conditions under which the fixed point components admit a cell decomposition.
Here the Caldero-Chapoton map comes into play.
In the case of the generalized Kronecker quiver, it turns out that a natural filtration of a fixed preprojective representation by preprojectives of smaller dimension transfers to the universal covering quiver.
These filtration can be successively described by short exact sequences.
The main advantage when passing to the universal covering is that the preprojective representations covering the same dimension vector below become orthogonal, a property which rigidifies the situation in a sense.
In the end, this machinery can be used to recursively build cell decompositions of all quiver Grassmannians of lifted (truncated) preprojective representations.
As all the quiver Grassmannians of the (non-lifted) representation are smooth, this combines with the torus action to give a cell decomposition of these quiver Grassmannians.

@@

As mentioned above, the Laurent polynomial expressions for cluster variables have been described using the representation theory of quivers \cite{cc,ck}: the cluster variables are generating functions for Euler characteristics of quiver Grassmannians.
For rank two cluster algebras, the Laurent expressions of cluster variables can also be computed using a certain Dyck path combinatorics \cite{llz}.
The confluence of these results gives rise to a combinatorial construction for the Euler characteristics and counting polynomials of certain quiver Grassmannians \cite{rupel}.
A consequence of our main result is a geometric explanation for these computations: we provide a one-to-one correspondence between the strong successor closed subsets and compatible pairs for an appropriate Dyck path which leads to a geometric explanation for the appearance of Dyck path combinatorics in the theory of quiver Grassmannians. 
\sayD{I need to include a conjecture about the dimensions of cells coming from the combinatorial construction of counting/Poincare polynomials}
